Ludwig User Tutorial

### 2017-06-16 docs/tutorial/tutorial.tex
Added test 1 to ludwig tutorial
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@3135 40876441-3496-0410-a613-a2fca35b2ae2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  ludwigTutorial.tex
%
%  Introductory tutorial for new users of Ludwig. 
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre and
%  Department of Physics, University of Strathclyde
%
%  Contributing authors:
%  Oliver Henrich (oliver.henrich@strath.ac.uk)
%  Fraser Mackay (s1026487@sms.ed.ac.uk)
%
%  (c) 2008-2017 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\usepackage{amsmath}
\usepackage{moreverb}
\usepackage{lscape}
\usepackage{epic}
\usepackage[pdftex]{graphicx}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{color}
\usepackage{listings}
\usepackage{mathtools}
\usepackage{enumerate}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  tutorial.tex
%
%  Introductory tutorial for new users of Ludwig. 
%
%  Edinburgh Soft Matter and Statistical Physics Group and
%  Edinburgh Parallel Computing Centre and
%  Department of Physics, University of Strathclyde
%
%  Contributing authors:
%  Oliver Henrich (oliver.henrich@strath.ac.uk)
%  Fraser Mackay (s1026487@sms.ed.ac.uk)
%
%  (c) 2008-2017 The University of Edinburgh
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\usepackage{amsmath}
\usepackage{moreverb}
\usepackage{lscape}
\usepackage{epic}
\usepackage[pdftex]{graphicx}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{color}
\usepackage{listings}
\usepackage{mathtools}
\usepackage{enumerate}
\graphicspath{{./tutorialFigs/}}

@@

\section{Tutorial Test Exercises}

\subsection{Test 1: Poiseuille Flow of a Simple Fluid in a Cavity}

In this section, the first test exercise example, the Poiseuille flow of a simple fluid in a cavity, is outlined. This example gives a step-by-step guide example for performing a short run ($10,000$ time-steps) of the \texttt{Ludwig} code for a simple system and provides a short introduction to the post-processing and data analysis procedure. Before starting, the code should be compiled in parallel and the executable \texttt{Ludwig.exe} should be copied to your working directory.

\begin{enumerate}
\item Copy the file \texttt{input} located in the directory \texttt{ludwig/docs/tutorial/test1} to your working directory. Note that this input file is a modified version of the reference input file \texttt{input.ref} located in \texttt{/src} containing the required \com{variables??} for this run. 
\item Run the code in your working directory by issuing the command:
\begin{lstlisting}
$ mpirun -np 4 ./Ludwig.exe input
\end{lstlisting}
\item You should now see output files with names of the form \texttt{vel-XXXXXXXX.001-001} where \texttt{XXXXXXXX} is replaced by the time-step number in increments of $1000$. These files contain the components of the velocity vectors in Cartesian coordinates at each lattice point in the simulation. 
\item The first step to the post-processing of these data is to copy the files \texttt{vtk\_extract.c} and \texttt{vtk\_extract\_script.py} from \texttt{/trunk/util} into your working directory.
\item Edit \texttt{vtk\_extract.c}. Specifically, set the flags for:
\begin{enumerate}
\item \texttt{input\_binary\_=0} \\ i.e. The data in the velocity files (\texttt{vel-XXXXXXXX.001-001}) are in ASCII format.
\item \texttt{output\_binary\_=0} \\ i.e. The output should also be in ASCII format.
\item \texttt{output\_index=1} \\ i.e. Produce the output data with lattice identifiers in the text files.
\item \texttt{vtk\_header=0} \\ i.e. The data files should not have column headers (these will be required later for visualisation in \textit{ParaView})
\item \texttt{output\_cmf=0} \\ i.e. The data should not be in column-major format (also required for \textit{ParaView}).
\end{enumerate}
\item Edit \texttt{vtk\_extract\_script.py}. Set:
\begin{enumerate}
\item \texttt{nstart=1000}, \texttt{nint=1000} and \texttt{nend=10000} to set the range and increments of the velocity data.
\item \texttt{vel=1} for velocity post-processing. 
\end{enumerate}
\item You should now see files named \texttt{vel-XXXXXXXX.vtk}. The first three columns of these text files are the spacial lattice indices followed by the components of the velocity vectors in Cartesian coordinates. To plot the velocity profile of the system, you can plot the $y$-component of the velocity against the spacial $x$-coordinate on the lattice with the command:
\begin{lstlisting}
gnuplot> p`vel-XXXXXXXX.vtk' u 1:5
\end{lstlisting} 
By plotting the data in each file, you should be able to see the velocity profile converging to the final profile as shown in Fig. \ref{fig:velocityProfile}. For reference, an reference example of the final output at $10,000$ time-steps can be found, in addition to a file containing the standard terminal output of the code in the directory \texttt{/trunk/docs/tutorial/test1/}.
\end{enumerate}

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.6\linewidth]{velProf.png}
  \caption{Plot of the parabolic velocity profile of a simple fluid in a cavity after $10,000$ time-steps}
  \label{fig:velocityProfile}
  \end{center}
\end{figure}
